\chapter*{Curriculum Vitae}\label{ch: curriculum}
\addcontentsline{toc}{chapter}{\nameref{ch: curriculum}}
\markboth{CURRICULUM VITAE}{CURRICULUM VITAE} % otherwise it still says "bibliography" on top due to chapter*

\vspace{-20pt}
\section*{Niels Claes}
\subsection*{Personal information}
\begin{tabular}[t]{ll}
  Date of birth & 21 April 1994  \\
  Nationality   & Belgian \\
  Email         & \href{mailto:niels.claes@live.be}{niels.claes@live.be} \\
  LinkedIn      & \href{https://www.linkedin.com/in/niels-claes-611140123/}{Niels Claes} \\
  ORCiD         & \href{https://orcid.org/0000-0002-8720-9119}{0000-0002-8720-9119}
\end{tabular}

\subsection*{Education}
\begin{itemize}
  \item[] \textbf{Doctor of Science (PhD) in mathematics} \hfill \\
  2018-2022, Centre for mathematical Plasma-Astrophysics, KU Leuven, Belgium \hfill \\
  Supervisor: Prof. Dr. Rony Keppens \hfill \\
  Thesis title: \emph{An MHD eigenspectrum approach to thermal instabilities}

  \item[] \textbf{Master of Science (MSc) in astronomy} \hfill \\
  2016-2018, KU Leuven, Belgium \hfill \\
  Supervisor: Prof. Dr. Rony Keppens \hfill \\
  Thesis title: \emph{Solar prominence formation and internal dynamics}

  \item[] \textbf{Bachelor of Science (BSc) in physics} \hfill \\
  2012-2016, KU Leuven, Belgium \hfill \\
  Supervisor: Prof. Dr. Christ Glorieux \hfill \\
  Thesis title: \emph{Investigating the limits of echolocation recognition}
\end{itemize}



\subsection*{Teaching experience}
\begin{itemize}
  \item[] \textbf{Calculus I} \hfill \\
  2018-2022 (bachelor in physics, mathematics), KU Leuven, teaching assistant.

  \item[] \textbf{Calculus II} \hfill \\
  2018-2022 (bachelor in physics, mathematics), KU Leuven, teaching assistant.

  \item[] \textbf{Computational methods for astrophysical applications} \hfill \\
  2018-2019 (master in astronomy), KU Leuven, teaching assistant.

  \item[] \textbf{Wiskunde II} \hfill \\
  2018-2019 (bachelor in geography), KU Leuven, teaching assistant.

  \item[] \textbf{Calculus \& analysis} \hfill \\
  2019-2022 (bachelor in bio-engineering), KU Leuven, teaching assistant.
\end{itemize}



\subsection*{Publications}
\begin{itemize}
  \item[] De Jonghe, J., \textbf{Claes, N.}, Keppens, R.: 2022,
  \emph{Legolas: magnetohydrodynamic spectroscopy with viscosity and Hall currents}.
  In preparation.

  \item[] \textbf{Claes, N.}, Keppens, R.: 2021,
  \emph{Magnetohydrodynamic spectroscopy of a non-adiabatic solar atmosphere}.
  Solar Physics, 296, 143 (24p).

  \item[] \textbf{Claes, N.}, De Jonghe, J., Keppens, R.: 2020,
  \emph{Legolas: A modern tool for magnetohydrodynamic spectroscopy}.
  Astrophysical Journal Supplement Series, 251, 25 (28p).

  \item[] \textbf{Claes, N.}, Keppens, R., Xia, C.: 2020,
  \emph{Thermal instabilities: Fragmentation and field misalignment of filament fine structure}.
  Astronomy \& Astrophysics, 636, A112 (12p).

  \item[] \textbf{Claes, N.}, Keppens, R.: 2019,
  \emph{Thermal stability of magnetohydrodynamic modes in homogeneous plasmas}.
  Astronomy \& Astrophysics, 626, A96 (14p).

  \item[] \textbf{Claes, N.}, Zelem, L., Pelegrin Garcia, D., Rychtarikova, M., Glorieux, C.: 2016.
  \emph{Echolocation recognition of different wall textures}.
  In: Proceedings of EAA EuroRegio 2016, (Paper No. 51), (1-9).
\end{itemize}

\subsection*{Colloquia and talks}
\begin{itemize}
  \item[] Talk (online), AGU 2021, New Orleans, USA, December 2021
  \item[] Talk and workshop, Magnetohydrodynamics of the 21st century, Leiden University, The Netherlands, October 2021
  \item[] Seminar, CmPA, KU Leuven, Belgium, October 2021
  \item[] Talk (online), AAPPS-DPP 2021, Japan, September 2021
  \item[] Talk (online), COSPAR 2021, Sydney, Australia, February 2021
  \item[] Talk (online), MHD Coronal Seismology meeting, University of Warwick, UK, December 2020
  \item[] Talk (online), AAPPS-DPP 2020, Japan, October 2020
  \item[] Solar physics seminar (invited), Northumbria University, UK, September 2020
  \item[] Seminar, CmPA, KU Leuven, Belgium, December 2019
  \item[] Talk and research visit, School of Astronomy and Space Science, Nanjing University, China, April-June 2019
  \item[] Seminar, CmPA, KU Leuven, Belgium, January 2019
\end{itemize}









\cleardoublepage
